Qualifying Report

### 2019-01-05 Source/QualifyingReport/QualifyingReport.tex
test
Added extra to QR

Despite this growing problem, at present there are no drugs that improve the prognosis of those suffering with AD. All the drugs that are on the market are designed to manage symptoms. Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Alzheimer2016}. Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place.

@@

@@

Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent. Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}. The case of Iris Murdoch illustrates this theory well. Le et al \cite{Le2012}found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel. If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed. However, one of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia.

@@

Many researchers have studied the early detection of AD. These studies usually follow two main approaches. The analysis of biomarkers and the examination of patients who have demonstrated decreasing cognitive abilities. The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease. The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD. In addition, the analysis of the decline of cognitive abilities is comparitively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved.
/par

@@

The consensus is that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals. 

@@

One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan. They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease. It turned out later to be the case that he had Alzheimer's Disease. Another classical study by Snowdon et al (1996) \cite{Snowdon1996} looked at whether linguistic ability in early life was associated with cognitive function and AD in later life. They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life. They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life. As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses.

@@

A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science. The following results were found (Table 1). All papers were then reviewed for relevance by reading the abstract and full text where appropriate and a shortlist was compiled. An additional search through references of shortlisted papers was also conducted and any papers who upon further review appeared relevant were added to the shortlist. Papers were included where researchers used machine learning to classify participants as MCI, AD or Healthy using language. We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English. This resulted in 17 journal articles and/or conference papers which form this review.

@@

In terms of standardised cognitive tests, there are two main aims. Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and, does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility). This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their perfomance in terms of both diagnostic and prognostic utility. \newline

@@

\subsubsection{RBANS - Repeated Battery for the Assessment of Neuropsychological Status}
\subsubsection{Patient Health Questionnaire)}
\subsubsection{WASI}
\subsubsection{Digit Span}
\subsubsection{Rey Words}
\subsubsection{Prospective Memory Recall}
\subsubsection{Digit symbol coding}
\subsubsection{Corsi Span}
\subsubsection{Verbal Fluency}
%% In this paragraph cite Henry et al, 2004 - taken from Taler and Phillips (2008) - Henry et al, 2004 is a meta analysis
There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories. Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter. Category fluency involves the generation of as many words as possible that fall into a specfied category. Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses. However, these two tasks require different strategies when attempting them. Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully. \par  
There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories. A review carried out by Henry et al, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
\subsubsection{Rey Delayed Recall}
\subsubsection{Naming - Speeded}

\subsubsection{Wisconsin Card Sorting}
\subsubsection{Complex Figure Task immediate and delayed recall}
\subsubsection{Cinderella Story - Narrative Description}
\subsubsection{Sentence Completion}
\subsubsection{Non-word Repetition}
\subsubsection{Grooved Peg Board}
\subsubsection{Visual Search}
\subsubsection{Spelling}
\subsubsection{Continuous Picture Naming}
\subsubsection{Free Cued Selective Reminding Test}
The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition. Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cogntive decline due to a cognitive impairment. /newline
/par
FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance (Sarazin et al, 2007 /newline
/par
The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score. /newline
/par

\subsection {Types of Language Assessment}

@@

>>>>

Despite this growing problem, at present there are no drugs that improve the prognosis of those suffering with AD. All the drugs that are on the market are designed to manage symptoms. Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Alzheimer2016}. Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place. It is therefore important to focus AD at it's earliest stages which some literature describes as 'Mild Cognitive Impairment (MCI) due to AD'.

@@

Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent. Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}. The case of Iris Murdoch illustrates this theory well. Le et al \cite{Le2012}found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel. If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed. However, one of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of dementia. Albert et al have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline. They note that the diagnosis of MCI requires evidence of intraindividual change and optimally requires evaluation at two or more points.
\begin{enumerate}
	\item Concern regarding a change in cognition - A person or an informant should express concern that there is a change in cognitive ability in comparison to previous level of performance.
	\item Impairment in one or more cognitive domains - There should be evidence of lower performance in one or more cognitive domains beyond what would be expected of a person given their age and education. 
	\item Preservation of independence in functional - Whilst persons with MCI are expected to be able to maintain independence, it is common to experience mild problems in complex functional tasks which they may have been able to perform previously. This might mean that they take more time or be less efficient at completing these tasks, or it may be that may make more mistakes.
	\item Not demented - The deterioration should be mild to the point that there is no significant loss of functioning in social or occupational contexts.
\end{enumerate}
In addition to meeting the above criteria, a clinician must rule out other conditions or factors that could account for the decline in cognition with the goal to increase the likelihood that the underlying cause of this decline is AD. 

@@

Many researchers have studied the early detection of MCI \ AD. These studies usually follow two main approaches. The analysis of biomarkers and the examination of patients who have demonstrated decreasing cognitive abilities. The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease. The second approach has gained more attention in recent years, due to the fact that in clinical practice it has shown promise in the early detection of AD. In addition, the analysis of the decline of cognitive abilities is comparitively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved. 

@@

@@

@@

One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan. They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease. It turned out later to be the case that he had Alzheimer's Disease. Another classical study by Snowdon et al (1996) \cite{Snowdon1996} looked at whether linguistic ability in early life was associated with cognitive function and AD in later life. They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life. They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life. As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses. The consensus among researchers that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals. 

@@

A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science. The following results were found (Table 1). All papers were then reviewed for relevance by reading the abstract and full text where appropriate and a shortlist was compiled. An additional search through references of shortlisted papers was also conducted and any papers who upon further review appeared relevant were added to the shortlist. Papers were included where researchers used machine learning to classify participants as MCI, AD or Healthy using language. We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English. 

@@

In terms of standardised cognitive tests, there are two main aims. Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and, does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility). This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their perfomance in terms of both diagnostic and prognostic utility.

@@

\subsubsection{Repeatable Battery for the Assessment of Neuropsychological Status}
The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specfically looking at detecting a characterizing very mild dementia. The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests. They met a number of design goals for this new battery of tests that addressed these shortcomings. The RBANS consists of a number of sub-tests across five distinct domains.

\begin{enumerate}
	\item Immediate Memory 
	\item Visuospatial / Constructional
	\item Language
	\item Attention
	\item Delayed Memory 
\end{enumerate}

However, whilst the authors claim that the RBANS is adequately sensitive in person's with MCI, other research points out that the RBANS has poor sensitivity in detecting MCI. Another drawback of this battery is the lack of executive function measures and object naming tasks. Research has shown that the RBANS has good test-retest reliability and convergent validity, pa

\subsubsection{Patient Health Questionnaire)}
The Patient Health Questionnaire
\subsubsection{WASI}
\subsubsection{Digit Span}
\subsubsection{Rey Words}
\subsubsection{Prospective Memory Recall}
\subsubsection{Digit symbol coding}
\subsubsection{Corsi Block Tapping Task}
The original Corisi apparatus consists of a series of nine blocks arranged irregularly on a 9 x 11in board. The blocks are tapped by the adminstrator in a randomized sequence of increasing length. Immediately after each sequence is demonstrated, the task is to mirror the sequence back to the adminstrator with the sequences increasing in length until a mistake is made. 
\subsubsection{Verbal Fluency}
%% In this paragraph cite Henry et al, 2004 - taken from Taler and Phillips (2008) - Henry et al, 2004 is a meta analysis
There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories. Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter. Category fluency involves the generation of as many words as possible that fall into a specfied category. Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses. However, these two tasks require different strategies when attempting them. Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully. \par  
There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories. A review carried out by Henry et al, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
\subsubsection{Rey Delayed Recall}
\subsubsection{Naming Tests}
Word-finding difficulty is a common symptom of AD and these deficits usually occur during the early stages of the disease progression. As such , a test of a patients ability to find words (known as confrontation naming) is a common way to measure cognitive decline. One common way to do this is the Boston Naming Test (BNT) which comprises 60 items on a spectrum of very frequent to very infrequent. This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span.

Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
\subsubsection{Wisconsin Card Sorting}
\subsubsection{Complex Figure Task immediate and delayed recall}
\subsubsection{Cinderella Story - Narrative Description}
\subsubsection{Sentence Completion}
\subsubsection{Non-word Repetition}
\subsubsection{Grooved Peg Board}
\subsubsection{Visual Search}
\subsubsection{Spelling}
\subsubsection{Continuous Picture Naming}
\subsubsection{Free Cued Selective Reminding Test}
The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition. Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cogntive decline due to a cognitive impairment. /newline
/par
FCSRT-IR (The instant recall version of the Free Cued Selective Reminding Test) performance has beeen shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the authors to define prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT-IR performance (Sarazin et al, 2007 /newline
/par
The sensitivity and specificity of the different test forms in correctly classifying patients were compared to each other using a previously established FCSRT-IR cut off score. /newline
\subsubsection{Conclusion}
I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population. However, particularly with the RBANS, there lacks sufficient sensitivity in differentiating those with MCI from healthy elderly individuals such that this tool could be used with a level of confidence in the results. In regard to the tests, a common theme is a lack of studies and therefore evidence into the utility of these tests with this particular population. However if, as researcher, we aim to investigate this population further then a benchmark battery of tests which is sensitive enough in both diagnostic and/or prognostic utility should be a goal.

\subsection{Types of Language Assessment}

@@

\subsubsection{How can Machine Learning be applied to this problem}
